\documentclass[12,twoside]{article}
\usepackage{url}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{framed}
\lstset{language=Python, keywordstyle=\color{blue}\bfseries, }
\usepackage{amsmath}

\pagestyle{myheadings}

\bibliographystyle{siam}

\addtolength{\textwidth}{1.00in}
\addtolength{\textheight}{1.00in}
\addtolength{\evensidemargin}{-1.00in}
\addtolength{\oddsidemargin}{-0.00in}
\addtolength{\topmargin}{-.50in}

\hyphenation{in-de-pen-dent}


\title{\textbf{Progress Report: Designing an Interpreter for a dynamic language 
  for graph algorithms}}

\author{Sandeep Dasgupta\thanks{Electronic address:
\texttt{sdasgup3@illinois.edu}}}

\begin{document}
\begin{titlepage}
\thispagestyle{empty}
\maketitle
\pagebreak
\end{titlepage}

\section{Problem Statement}

   In this project we are planning to work on a dynamically typed language to 
   represent graphs and apply various computations on them.
 
\section{Language Syntax}
  The syntax of the language is an oversimplified version  of C, but
    without mention of any types. The operations on incompatible types
    will be error-ed out while interpreting.

  \subsection{Progress}
  \begin{itemize}
    \item We have implemented the tokenizer and syntax analyzer using flex and 
    bison.  
    \item We are supporting syntax like \#include("filename") and 
    \#define("PI", "3.14") while doing a single pass of parsing  (i.e. 
        Preprocessing of these constructs are done while parsing). This is 
    achieved by using flex's internal stack to manage multiple buffers.

    \item The gramer rules are borrowed from 
    ~\url{http://www.quut.com/c/ANSI-C-grammar-y.html}. The rules are compiled 
    by bison tool to generate the C parser. 

    \item  We are able to generate the AST corresponding to the grammer rule for 
    a confirming test case. Our AST is basically a list of 
    function definitions. Each function definition class contains name of the 
    function, a set of formal arguments and a list of body statements. These 
    body statements could be a assignment, loop-statement, function call, etc.
    The leafs of the AST could be an identifier, int, float, true, false, null, 
    string, vertex, edge or graph.

    \item Some of the key features of the parser is as follows:
    \begin{itemize}
      \item Support of C statements like  {\tt\emph{if then}}, {\tt\emph{if then 
        else}}, {\tt\emph{while}}, {\tt\emph{for}}. 

      \item Support of {\tt\emph{break}}, {\tt\emph{continue}} within loop-body 
      and {\tt\emph{return}} in function-body.
      As we are representing both loop-body and function-body (i.e. anything 
          between ``\{'' \& ``\}'') as compound statements so we do not have to
      distinguish these two cases. The semantics of executing a 
{\tt\emph{break}}, {\tt\emph{continue}} and {\tt\emph{return}} will be discussed 
in the interpreter runtime section.

      \item Supporting graph as first class object {\tt\emph{valuegraph}}.
      The syntax to declare a graph is {\tt\emph{g = graph();}} which will be 
      represented in AST as an assignment-node with left-node containing an 
      identifier and right
      node as a function call. Now this function call corresponds to a built in 
      function that returns a {\tt\emph{valuegraph}} (which is of one the leaf 
          nodes of AST) on execution.

      \item Supporting vertices and edges as first class objects which contains 
      a map to add properties. This feature is useful in various graph 
      algorithms like in dfs traversal we may use a vertex property ``visited'' 
      to keep track of vertices already explored.
    \end{itemize}
  \end{itemize}
  


\section{Language Semantics}
The language semantics will be same as that of C as we are using a subset of it.

\section{Interpreter Runtime}
The following are the key features of the runtime:
\begin{itemize}

  \item The runtime starts with searching for function definition
  {\tt\emph{function main(argv)}} and then creats a function call out of it and 
  execute it.  While creating the function call it uses the command-line 
  parameters as the actual parameters of the function call.

  \item The execution of a function call involves finding the correspoding 
  function definition, checking if the number of formal and actuals are equal 
  and then pushing a call stack frame ( which contains the mapping between the 
      formal and actual values passed to them) in a global call stack. After 
  that, the function
  is executed w.r.t the current context(i.e. the top of the call stack).

  \item The execution of the function involves executing a list of statements.
  The statements may add further mappings in the current call stack frame.
  Whenever a name (identifier) is refereed, the mapping in the current context 
  need to be consulted to get the actual value of it. 

  \item The semantics of {\tt\emph{break}}, {\tt\emph{continue}} or 
{\tt\emph{return}} is supported using the try-catch mechanism of C++.
  For example, while interpreting a {\tt\emph{loop-body}}, whenever a {\tt\emph{break}}is 
  encountered, a corresponding {\tt\emph{break-object}} is thrown, which is caught in a place 
  outside the entire loop execution in order to implement the semantics of break. 

  Similarly, while executing a {\tt\emph{loop-body}}, whenever a 
{\tt\emph{continue}} is encountered, a {\tt\emph{continue-object}} is thrown, which is caught outside
  the {\tt\emph{loop-body}} execution so as to skip the current iteration 
    and continue with the {\tt\emph{loop-incr}} execution (in case of for loop) 
    and {\tt\emph{loop-condition-expr}} execution.

  And finally, while executing a  {\tt\emph{function-body}}, which is a list
  of statements, when any one of those statements is a {\tt\emph{return}}, a 
  {\tt\emph{return-object}} is thrown, which is caught outside of the loop
  which is going over that list and in this way the semantics of return is 
    maintained.
  
  \item Occurrence of {\tt\emph{break}} and {\tt\emph{continue}} within non-loop
    body triggers an error. While interpreting a node-block (which is a set 
        statements within ``\{'' and ``\}''), whenever the runtime finds a 
{\tt\emph{break}} or  {\tt\emph{continue}} it throws a object. Now if this 
object is caught
  inside a non-loop block then error is reported.

  \item Division by zero and operations on incompatible types are runtime 
  errors.  
  
  \end{itemize}

\section{Future Work}
The following are the future work.
  \begin{itemize}
    \item To support  additional operators to specify relationships between the 
    nodes/edges or their groups. The operator that we are planning to support 
    are the described next with their semantics.

    \begin{itemize}
      \item v1$\leftarrow$v2: Select all the directed edges from vertices v1 to 
      v2.
      \item v1$\rightarrow$v2: Select all the directed edges from vertices v2 to 
      v1.
      \item v1$\leftrightarrow$v2: Select all the directed edges between 
      vertices v1 and v1.
      \item v1$?$v2: Select all the edges (directed or undirected) between 
      vertices v1 and v1.
      \item S1$\cap$S2: Selects the intersection between the set of
      vertices/edges.
    \end{itemize}
      Most of these ideas behind choosing the operators are borrowed from 
      Guess~\cite{Adar}. The intuition behind these operators is that the users do 
      not have to remember longer commands. Also its very intuitive to build up 
      complex operations using the simpler ones.

    \item To support saving of state and retrieving it back using
    routines like saveSateFromFile \& loadStaeFromFile.
    This idea is borrowed from Guess~\cite{Adar} and this seems a useful service 
    provided to do exploratory programming, as the user might be interested to 
    checkout the last saved state  (or a state with a any tag) or to undo all 
    the experiments down to a particular state.

  \end{itemize}

\subsection{Evaluation Strategy}
The baseline of our evaluation will be the open source Graphal~\cite{Graphal} system.  
Graphal~\cite{Graphal} is an interpreter of a programming language that is mainly 
oriented to graph algorithms. There is a command line interpreter and a 
graphical integrated development environment. The IDE contains text editor for 
programmers, compilation and script output, advanced debugger and visualization 
window. The evaluation will be between command line version of Graphal with our
implementation.
  The evaluation will
  cover the following two aspects of our implementation:
  \begin{itemize}
    \item The convenience and intuitive extensibility provided by our 
    programming model.
    \begin{itemize}
      \item We will be implementing a couple of well know graph algorithms in 
      both baseline and our implementation
        and use number of dynamic instructions interpreted as a measure to show 
        the conciseness of our
        representation.
    \end{itemize}
    \item Performance
    \begin{itemize}
      \item As we are planning to keep the compiled version of frequently used 
      graph routines (like dfs\_iterators(vertex), bfs\_iterators(vertex)), we are 
      expecting to achieve better performance in terms of runtime.  
      \end{itemize}
  \end{itemize}





\bibliography{ProgressReport.bib}

%\nocite{*}


\end{document}
