In order to evaluate the performance of our design, we will be coding a number 
of well known graph related algorithms in our language and run the proposed interpreter on 
them. The baseline of our evaluation will the the same algorithms implemented in 
C,  compiled by the C compiler and executing the compiled binary.

We are planning to compare the following.
\begin{itemize}
  \item Runtime performance of the interpreter 
w.r.t the compiled version of the code.
  \item Convenience of representation in our 
language w.r.t C language.
\end{itemize}


\subsection{Runtime Performance Comparison}

We have implemented a couple of algorithms in both our language and C. The 
algorithms are Transitive Closure of a graph using Floyd Warshall's algorithm, 
           Shortest Path using Dijkstra's algorithm, Minimum Spanning Tree using 
           Prim's algorithm and Chaitin's Optimistic graph coloring algorithm.
The input graph used in all the cases is a randomly generated graph consisting  of vertices = 500, edges = 1000. 
For Graph Coloring, we used the input graph with vertices = 125, edges = 1000. This is 
because it takes a lot of time for our interpreter execution to complete.
In Table~\ref{table:eval_1}, we compared the runtime of our interpreter 
(henceforth called GRI) with 
execution time of compiled binary obtained by compiling the C implementation.

Note that, while implementing the algorithms, we used the basic syntax provided 
by our implementation and \textbf{NOT} used any of the  built-in 
functions like {\tt getTransitiveClosure()}, {\tt getShortestPath()} or {\tt getMST()}. 
We call such implementations as fully scripted.

  \begin{table}[]
\centering
\caption{Slowdown of GRI w.r.t C implementation.}
\label{table:eval_1}
\scalebox{.8} {
\begin{tabular}{|l|l|l|l|}
\hline
   \multicolumn{1}{|c|}{Algorithm}                                 & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Fully Scripted\\ time (secs)\end{tabular}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}C implementation\\ time(secs)\end{tabular}} & Slowdown \\ \hline
Transitive Closure  & 1137.40            & 4.04                   & 281.5    \\ \hline
Shortest Path          & 6.37             & 0.02                     & 318.5     \\ \hline
Minimum Spanning Tree        & 6.34             & 0.02                      & 317.0     \\ \hline
Graph Coloring  &   138.17  & 0.65                              & 212.6         \\ \hline
\end{tabular}}
\end{table}

As we can see that the slowdown are huge. To counter that, we planned to come up 
with following built-in functions.
  \begin{itemize}
    \item { \tt Graph::getAdjMatrix();}
    \item {\tt Graph::getTransitiveClosure();}
    \item {\tt Graph::getShortestPath(wt, start, end);}
    \begin{itemize}
      \item Returns an array of 2 elements. 
      \begin{itemize}
        \item First is an array representing the parent of each vertex in the 
        shortest path tree. 
        \item Second is another array representing the shortest distance from 
        vertex start to each vertex in the shortest path tree.
      \end{itemize}
      \item If {\tt end == NULL}, the the algorithm will compute the shortest
distance of all the vertices from start. Else it stops when the
shortest distance of end vertex from start is computed.
      \item Weight ``wt'' (a string) signifies which of the properties of the
edge need to be considered as distance in the shortest path algorithm.
    \end{itemize} 
    \item {\tt Graph::getMST(wt);}
    \begin{itemize}
      \item Return the parent of each vertex in the minimum spanning tree.
      \item Weight ``wt'' (a string) signifies which of the properties of the
edge need to be considered as distance in the minimum spanning tree algorithm.
    \end{itemize} 
  \end{itemize}


  Table~\ref{table:eval_2} shows the speedup of our implementation \textbf{with} built-in 
  functions w.r.t  \textbf{without} using built-in functions (i.e. fully 
      Scripted). The graph used in 
  all the cases consist of vertices = 125, edges = 1000.

\begin{table}[]
\centering
\caption{Speedup with built-in functions w.r.t without 
  using built-in functions.}
\label{table:eval_2}
\begin{tabular}{|l|l|l|l|}
\hline
       \multicolumn{1}{|c|}{Algorithm}                             & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}With Built-ins\\ time (secs)\end{tabular}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Fully Scripted\\ time(secs)\end{tabular}} & Speedup \\ \hline
Transitive Closure & 0.54                                                                                      & 18.19                                                                                    & 33.6    \\ \hline
Shortest Path             & 0.08                                                                                      & 0.51                                                                                     & 6.3     \\ \hline
Minimum Spanning Tree        & 0.05                                                                                      & 0.47                                                                                     & 9.4     \\ \hline
\end{tabular}
\end{table}

Finally, we replaced the fully scripted version of the algorithms with built-in 
function calls and compared that with C (Figure ~\ref{table:eval_3}). 
The graph used in all the cases consist of vertices = 500, edges = 1000. 
For Graph Coloring, we used the graph with vertices = 125, edges = 1000.

  \begin{table}[]
\centering
\caption{Slowdown of our implementation w.r.t C implementation.}
\label{table:eval_3}
\begin{tabular}{|l|l|l|l|}
\hline
         \multicolumn{1}{|c|}{Algorithm}                           & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}GRI\\ Time(secs)\end{tabular}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}C\\ Time(secs)\end{tabular}} & Slowdown \\ \hline
Transitive Closure  & 10.26     & 4.04                               & 2.5     \\ \hline
Shortest Path            & 0.09      & 0.02                               & 4.5     \\ \hline
Minimum Spanning Tree        & 0.09      & 0.02                               & 4.5     \\ \hline
Graph Coloring  &   138.17  & 0.65                              & 212.6         \\ \hline
\end{tabular}
\end{table}

In Graph Coloring algorithm, we are still loosing a lot on performance due to 
the fact that the implementation is fully scripted and its interpretation 
is very slow as compared to binary execution.

\subsection{Comparison w.r.t Representation Convenience }
We will be using the number of lines of code as a metric to compare the 
convenience that our language is providing w.r.t C. 

As we are aware of the fact that this metric is not that useful if we are 
comparing two languages having different conventions for how much we should put on a line. 
But as our language is a subset of C, so  this metric makes sense, but 
to yield 
meaningful results we need to  be careful about the following aspects.
\begin{itemize}
  \item Both the code implementations should produce the same output.
  \item In both the code implementation, the backward slice of the output instructions should cover the entire 
  code. In other words, there should not be any redundant code which is not 
  contributing to the generated output. 
  \item Both the code implementation should follow similar coding 
  guidelines~\cite{GNU}.
  \item Both the programs need to be equally readable in terms of indentation, 
  newlines, spaces and comments.
\end{itemize}

 We compared the number of lines of static code written in our language with 
 that written in C.  Table~\ref{table:eval_4} represents the results of this 
 comparison, which clearly shows that we need fewer lines of code to implement 
 the same algorithm.

 \begin{table}[h]
\centering
\caption{Comparison of Number of lines of code in our implementation w.r.t C implementation.}
\label{table:eval_4}
\begin{tabular}{|l|c|c|}
\hline
\multicolumn{1}{|c|}{Algorithm} & Proposed language & C   \\ \hline
Transitive Closure (with built-in)              & 36                & 129 \\ \hline
Transitive Closure (fully scripted)              & 49                & 129 \\ \hline
Shortest Path (with built-in)                  & 64                & 191 \\ \hline
Shortest Path (fully scripted)                  & 159                & 191 \\ \hline
Minimum Spanning Tree (with built-in)           & 40                & 171 \\ \hline
Minimum Spanning Tree (fully scripted)          & 128                & 171 \\ \hline
Graph Coloring                  & 135               & 213 \\ \hline
\end{tabular}
\end{table}
