 The syntax of the language is an oversimplified version  of C, but
    without mention of any types. The operations on incompatible types
    will be error-ed out while interpreting.

    We have implemented the tokenizer and syntax analyzer using flex and 
    bison. We are supporting syntax like \#include("filename") and 
    \#define("PI", "3.14") while doing a single pass of parsing  (i.e.  
        Preprocessing of these constructs are done while parsing). This is 
    achieved by using flex's internal stack to manage multiple buffers.
    The grammar rules are mostly borrowed from 
    ~\cite{ANSI}. The rules are compiled 
    by bison tool to generate the C parser. 
    We are able to generate the AST corresponding to test-cases 
    confirming to the grammar rules. Our AST is basically a list of function 
    definitions. Each function definition object contains name of the function, 
    a set of formal arguments and a list of body statements. These body 
    statements could be an assignment, loop-statement, function call, etc.
    The leafs of the AST could be an identifier, int, float, true, false, null, 
    string, vertex, edge or graph.

    Some of the key features of the parser is as follows:
    \begin{itemize}
      \item Support of C statements like  {\tt if then}, {\tt if then 
        else}, {\tt while}, {\tt for}, {\tt foreach}. 

      \item Support of {\tt break}, {\tt continue} within loop-body 
      and {\tt return} in function-body.
      As we are representing both loop-body and function-body as compound 
      statements (i.e. anything between ``\{'' \& ``\}''), so we do not have to
      distinguish these two cases. But we will error-out if break is used inside 
      non-loop body. The detailed semantics of executing a {\tt break},   
{\tt continue} and {\tt return} will be discussed in the interpreter 
runtime section.

      \item Supporting graph as first class object {\tt graphnode}.
      The syntax to declare a graph is {\tt g = graph();} which will be 
      represented in AST as an assignment-node with left-node containing an 
      identifier and right
      node as a function call. Now this function call corresponds to a built in 
      function that returns a {\tt graphnode} (which is of one the leaf 
          nodes of AST) on execution.

      \item Supporting vertices and edges as first class objects which contains 
      a map to add properties. This feature is useful in various graph 
      algorithms like in dfs traversal~\ref{fig:motiv_5}, we uses a vertex property ``visited'' 
      to keep track of vertices already explored.

      \item We are supporting composite data-structures like array, struct and 
      set and iterators on them. Figure~\ref{fig:language_1} represents a code 
      snippet representing some of the operation on these data-structures.

      \item The language semantics will be same as that of C as we are using a subset of it.`

      \item All variables are defined as local and are valid only it the scope of the current function (function not block).

      \item The language specify no constructs for variable declaration and type 
      specification. The interpreter uses some inner data types (like {\tt 
          null, Bool, int, float, string, array, struct, set, graph, vertex, 
          edge})  which can be dynamically changed with assign command.

    \end{itemize}
    \input{Figs/fig_language_1.tex}
  
